---
title: "Advanced Visualization Code"
author: "Sofia Perez"
date: "2026-03-06"
format: html
editor: visual
---

## Set up

```{r}
#| message: false
library(tidyverse) # general use
library(here) # file organization
library(janitor) # cleaning data frames
library(readxl) # reading excel files
library(ggpattern) # geom pattern customization
library(ggthemes) # extra plot themes
library(showtext) # to show custom text on plot
library(sysfonts) # to read in custom google font
library(GGally) # for more complex plots (correlation heatmap)
library(labelled) # for customizaing variable labels
font_add_google(name = "Bellefair", # Name of the font on the Google Fonts site
                family = "belle") # Name to call the font
showtext_auto() # automatically use showtext to render text

personal_data <- read_csv(here("data",
                               "personal_data_03072026.csv")) # save personal data as object
personal_data_clean <- personal_data |> # create clean data frame of personal data
  clean_names() |> # clean column names to lowercase and underscores
  mutate(Date = as.Date.character(date, format = "%m/%d/%Y"), # mutate date column into readable dates 
         Game = time_with_mobile_game_active_min, # creating columns with updated names, new name = old name
         Class = time_in_class_lecture_hrs, #  new name = old name
         Study = time_spent_in_study_spaces_hrs, #  new name = old name
         Friends = time_spent_at_friends_hrs, #  new name = old name
         Assignments = number_of_assignments_due) |> #  new name = old name
  select(Date, Game, class, Class, Study, Friends, Assignments) # select columns of interest
```

# Visualization 1.

```{r}
#| label: hexplot class+game
#| fig-height: 6 # set custom figure height
#| fig-width: 6 # set custom figure width
hexplot <- ggplot(data = personal_data_clean, # create base ggplot layer with clean df
       aes(x = Class, # set x-axis
           y = Game)) + # set y-axis
  geom_hex(bins = 11) + # customize bin number  (to smallest # that won't cut off geom shape near border)
  scale_fill_continuous(type = "gradient", # customize color of hexbins
                        low = "#39240A", # set low count occurence color
                        high = "#A0A0A0") + # set higher count occurence color
  theme_void() + # set theme to remove axis and labels
  theme(legend.position = "none") + # remove legend
  theme(plot.margin = margin(t = 1, # create 1 cm margin around all sides
                             r = 1,
                             l = 1,
                             b = 1,
                             unit = "cm"),
         plot.background = element_rect(fill = "#F5F3EF")) + # set background color
  geom_curve(x = 2.25, # x-axis start point
             y = 50, # y-axis start point
             xend = 1.5, # x-axis end point
             yend = 45, # y-axis end point
             curvature = 0.6, # positive curve of 0.6 magnitude
             angle = 70, # angle where curve happens
             linewidth = 2, # customize width
             arrow = arrow(length = unit(0.5, "cm")), # add arrow to line and sest size
             lineend = "round", # customize end of line
             col = "#B4A073") + # set color
  annotate("text", # add in plot text annotation
           x = 2.53, # x coordinate for text
           y = 46, # y coordinate for text
           label = "Plot area divided into bins", # set text label
           color = "#39240A", # set color
           size = 12, # set size
           family = "belle") # set font
 
png("vizualization1_hexplot.png", # save visualization as png image
    bg = "transparent", # make background transparent
    width = 1192, # set width
    height = 1192, # set height
    res = 192, # set image resolution
    unit = "px") # use pixels as unit
plot(hexplot) # plot the visualization to save as png
dev.off() # finalize and close plot
```

# Visualization 2.

```{r}
#| label: correlation matrix
#| warning: false # remove warning messages when rendered
#| fig-height: 5 # custom figure height
#| fig-width: 5 # custom figure width
corrmatrix <- ggcorr(data = personal_data_clean, # create correlation heatmap with clean df
       method = c("pairwise", "spearman"), # choose method of correlation on data
       low = "#8500E9", # set negative correlation color
       mid = "#C0CDFF", # set close to 0 correlation color
       high = "#56E900") + # set positive correlation color
  theme_minimal(base_family = "belle") + # set theme to minimal and base text font
  theme(text = element_text(size = 30), # set font size
        plot.background = element_rect(fill = "#F5F3EF"), # set background color
        legend.position = "top") # put legend at top

png("vizualization2_corrmatrix.png", # save visualization as png image
    bg = "transparent", # make background transparent
    width = 1192, # set width
    height = 1192, # set height
    res = 192, # set image resolution
    unit = "px") # use pixels as unit
plot(corrmatrix) # plot the visualization to save as png
dev.off() # finalize and close plot
```

# Visualization 3.

```{r}
#| label: time series class+game
coeff <- 18 # define coefficient for later use

timeseries <- ggplot(data = personal_data_clean, # create base ggplot layer with clean df
       aes(x = Date)) + # set x-axis as date to create time series
  geom_line(aes(y = Class), # create line for time in class
            color = "#A0A0A0", # set color
            linewidth = 1) + # set width
  geom_line(aes(y = Game/coeff), # create line for time spent on game and divide values to scale it to existing y-axis values of hours
            color = "#0F0F0F", # set color
            linewidth = 1) + # set width
  scale_y_continuous(name = "Time in lecture (hrs)", # scale first y-axis to hours and name it and name it accordingly
                     sec.axis = sec_axis(~ . * coeff, # scale second y-axis to minutes and name it accordingly
                                         name = "Time tracked on game (min)")) +
  theme_classic() + # set theme to classic
  theme(text = element_text(family = "belle"), # set base font
        plot.background = element_rect(fill = "#F5F3EF"), # set entire background color
        panel.background = element_rect(fill = "#F5F3EF"), # set plot background color
        panel.border = element_rect(fill = NA, # create border, set color and width
                                    color = "#B4A073", 
                                    linewidth = 5),
        axis.title.x = element_text(size = 40, # set size and color of bottom axis
                                    color = "#39240A"),
        axis.title.y.left = element_text(size = 35, # set size and color of left axis
                                         color = "#A0A0A0"),
        axis.title.y.right = element_text(size = 35, # set size and color of right axis
                                          color = "#0F0F0F"),
        axis.text.x = element_text(size = 30, # set size and color of x-axis ticks text
                                   color = "#39240A"),
        axis.text.y = element_text(size = 35, # set size and color of y-axis ticks text
                                   color = "#39240A"),
        axis.line = element_blank() # remove axis lines
       )

png("vizualization3_timeseries.png", # save visualization as png image
    bg = "transparent", # make background transparent
    width = 1344, # set width
    height = 960, # set height
    res = 192, # set image resolution
    unit = "px") # use pixels as unit
plot(timeseries) # plot the visualization to save as png
dev.off() # finalize and close plot

print(timeseries)
```

### Infographic planning brainstorm notes

**Fonts** - Brothers, Sabbath Black

**Colors** - logo: \~gold "C6A06E" , white "FFFFFF" ;

Box border: \~wheat"B4A073" Box interior: \~cream "F5F3EF"

Stripe dark "C2B28E" Stripe light "CDC0A2" Perimeter \~black "0F0F0F"

Symbol accents: \~wheat" B4A073" \~black "0F0F0F" \~cream "F5F3EF" \~grey "A0A0A0

Font \~brown "39240A"

Shiny button \~Dgrey "333333" \~black "0F0F0F"

# Write-up

## General Information about design and visualization

patterns highlighting in each + why

outside of data viz, what aesthetic (color, font, arrangement) choices and why? What way do your choices contribute to a conmpelling/interesting narrative?

## Sources and process

for each viz describe examples that imspired that viz (cite author)

describe coding process, ex. "did you have to clean up/sammarize data before you started? Which geoms did you start with? What geoms did u end up using and why?

describe tools used: other people's code? Google/StackOverflow?
